### 2018-10-30 proposal.Rmd
proposal draft

---
title: "DS 5220 Project Proposal"
author: "Chaitya Shah, Tyler Brown, Karan Gulur Muralidhar, Omair Shafi Ahmed; MSDS 19'"
output: pdf_document
---

**Description of the problem**

Our team is basing our project idea on the "Quick, Draw!" Kaggle competition 
which classifies user drawings into one of 340 label categories. Drawings may 
be incomplete or not match the label. The challenge is to effectively build a 
recognizer that can work with this noisy data. The challenge is unique from an
algorithmic perspective because the data includes both time series and pattern
recognition components; areas that have traditionally used two different 
types of Neural Network Architectures. Our team uses a subset of the data to 
provide a solution to the Kaggle problem while taking a more in depth 
look at potential pitfalls and modeling strategy without having to use 
extensive computational resources.

**Summary of the data**

Google curated the Quick Draw Dataset as a collection of 50 million drawings
mapped to 345 categories. The drawings were submitted by players of the game
"Quick, Draw!" and captured as timestamped vectors, tagged with metadata such
as the requested drawing category and country code. The format of each drawing
array contains 3-dimensional vectors where components are the x,y coordinates
and time in milliseconds since the first point. We use a preprocessed version
of the dataset which is simplified and takes up about a tenth of the memory.
The simplified dataset contains a CSV file for each of the 345 categories.
Google was able to prepocess the data by performing the following steps: (1)
align each drawing to the top-left corner, (2) uniformly scale each drawing, 
(3) resample all strokes with a 1 pixel spacing, and (4) simplify all strokes
using the Ramer-Douglas-Peucker algorithm with an epsilon value of 2.0.

**Methods**

We plan to use a Neural Network for this problem. The problem is unique in
that it has a time series component related to drawing stoke sequence, and
a image recognition component related the cumulative strokes. Recurrent Neural
Networks (RNN) are generally associated with time series. Convolutional Neural
Networks (CNN) are often associated with image recognition. The inclusion of
time series and pattern recognition components requires us to consider 
tradeoffs related to model accuracy, complexity, and algorithmic scalability.
For example, RNNs have been shown to be less scalable than CNNs due to
computational resource requirements. We have noted that researchers have
attempted to capture benefits of RNNs in CNNs through the use of techniques
such as dilation. We are still working through the best Neural Network 
Architecture for this problem.

Feature engineering is also an important part of our modeling strategy. We
recognize that choosing appropriate features can reduce the amount of data 
required to train a Neural Network appropriately, as well as reduce the amount
of computational resources required to fit our model.

**Preliminary Results**

Given that Neural Networks can be computationally intensive, one of our
initial assessments was where and how to run these models. We tried 
replicating a CNN which had already worked for a team on Kaggle using the
Google Colab resources. We were unable to replicate their results. We then 
reached out to the Instructor to explore alternatives such as better 
computers or using a subset of our data. We chose to use a subset of the
data to focusing on modeling rather than computational infrastructure.

When visualizing our data, we were able to confirm that the data appears to
be very noisy. We found that "Quick, Draw!" players interpreted categories 
such as "cat" correctly but used different perspectives. One drawing may be 
the face of a cat, another included a side-view with the face, body, and tail.
Our model will need to account for these additional variations.

**References**

Image Based CNN (https://www.kaggle.com/jpmiller/image-based-cnn)

### 2018-12-01 report/report_draft.Rmd
added rough outline for the report

---
title: "Identifying Drawings with the Quick Draw Challenge"
author: "Blue Team"
output: 
  pdf_document:
    keep_tex: yes
bibliography: references.bib
---

We're trying to put together a CNN based on the "Quick, Draw!" Kaggle
competition [@QuickDra10].

# Theory & Literature Review

CNN with dilation is preferable to LSTM because it's presumably
faster. We can cite some papers here.

# Hypotheses

Maybe we should structure our models as experiments. Then talk 
about what worked the best. Here's some example hypotheses

$H_a:$ Prediction will go up when we use a CNN compared to a fully
	connected feedforward neural network.

$H_a:$ Prediction will go up when we use regularization

# Data

We got our data from Google/Kaggle

# Results

Here's how the results performed when we were exploring our 
hypotheses

# Discussion

Here's what worked the best, worst, and future areas for research.

# References

### 2018-12-01 report/report_draft.tex
added rough outline for the report

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Identifying Drawings with the Quick Draw Challenge},
            pdfauthor={Blue Team},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Identifying Drawings with the Quick Draw Challenge}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Blue Team}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \date{}
  \predate{}\postdate{}

\begin{document}
\maketitle

We're trying to put together a CNN based on the ``Quick, Draw!'' Kaggle
competition (Google 2018).

\hypertarget{theory-literature-review}{%
\section{Theory \& Literature Review}\label{theory-literature-review}}

CNN with dilation is preferable to LSTM because it's presumably faster.
We can cite some papers here.

\hypertarget{hypotheses}{%
\section{Hypotheses}\label{hypotheses}}

Maybe we should structure our models as experiments. Then talk about
what worked the best. Here's some example hypotheses

\(H_a:\) Prediction will go up when we use a CNN compared to a fully
connected feedforward neural network.

\(H_a:\) Prediction will go up when we use regularization

\hypertarget{data}{%
\section{Data}\label{data}}

We got our data from Google/Kaggle

\hypertarget{results}{%
\section{Results}\label{results}}

Here's how the results performed when we were exploring our hypotheses

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Here's what worked the best, worst, and future areas for research.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-QuickDra10}{}%
Google. 2018. ``Quick, Draw! Doodle Recognition Challenge \textbar{}
Kaggle.'' \url{https://www.kaggle.com/c/quickdraw-doodle-recognition}.

\end{document}

### 2018-12-01 report/report_draft.Rmd
update report

@@
 title: "Identifying Drawings with the Quick Draw Challenge"
-author: "Blue Team"
+author: "Group 3"
 output: 
@@
 
-# Theory & Literature Review
+# Introduction
 
-CNN with dilation is preferable to LSTM because it's presumably
-faster. We can cite some papers here.
+Our team has based our project idea on the "Quick, Draw!" Kaggle competition 
+which classifies user drawings into one of 345 label categories. Many drawings 
+were incomplete or failed to match the label. The challenge was to effectively 
+build a recognizer that could work with this noisy data. The challenge was unique 
+from an algorithmic perspective because the data included both temporal and 
+spatial components; areas that have traditionally used two different 
+types of Neural Network Architectures. Our team attempted to provide a solution 
+using a subset of the data, to the Kaggle problem, while taking a more in depth 
+look at potential pitfalls and modeling strategy, without having to use 
+extensive computational resources.
 
-# Hypotheses
+# Methods
+
+First, we got our data from Google/Kaggle. CNNs are traditionally known to work
+well with spatial data, which prompted us to process the original csv files to 
+images of the drawings.
 
@@
 
-$H_a:$ Prediction will go up when we use regularization
+$H_a:$ Prediction will go up when we use regularization.
 
-# Data
+$H_a:$ Run neural networks without any activations.
 
-We got our data from Google/Kaggle
+$H_a:$ Try different architectures- number of layers, different (or no) 
+activations
+
+$H_a:$ CNN with dilation is preferable to LSTM because it's presumably
+faster. We can cite some papers here.
 
@@
 # References
+
+# Appendix
+
+Add computer code, plots, and other relevant technical details that will help
+me evaluating your work.
+
+# Statement of contribution

### 2018-12-01 report/report_draft.tex
add report pdf

@@
             pdftitle={Identifying Drawings with the Quick Draw Challenge},
-            pdfauthor={Blue Team},
+            pdfauthor={Group 3},
             pdfborder={0 0 0},
@@
   \posttitle{\par}
-  \author{Blue Team}
+  \author{Group 3}
   \preauthor{\centering\large\emph}
@@
 
-\hypertarget{theory-literature-review}{%
-\section{Theory \& Literature Review}\label{theory-literature-review}}
+\section{Introduction}\label{introduction}
 
-CNN with dilation is preferable to LSTM because it's presumably faster.
-We can cite some papers here.
+Our team has based our project idea on the ``Quick, Draw!'' Kaggle
+competition which classifies user drawings into one of 345 label
+categories. Many drawings were incomplete or failed to match the label.
+The challenge was to effectively build a recognizer that could work with
+this noisy data. The challenge was unique from an algorithmic
+perspective because the data included both temporal and spatial
+components; areas that have traditionally used two different types of
+Neural Network Architectures. Our team attempted to provide a solution
+using a subset of the data, to the Kaggle problem, while taking a more
+in depth look at potential pitfalls and modeling strategy, without
+having to use extensive computational resources.
 
-\hypertarget{hypotheses}{%
-\section{Hypotheses}\label{hypotheses}}
+\section{Methods}\label{methods}
+
+First, we got our data from Google/Kaggle. CNNs are traditionally known
+to work well with spatial data, which prompted us to process the
+original csv files to images of the drawings.
 
@@
 
-\(H_a:\) Prediction will go up when we use regularization
+\(H_a:\) Prediction will go up when we use regularization.
 
-\hypertarget{data}{%
-\section{Data}\label{data}}
+\(H_a:\) Run neural networks without any activations.
 
-We got our data from Google/Kaggle
+\(H_a:\) Try different architectures- number of layers, different (or
+no) activations
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\(H_a:\) CNN with dilation is preferable to LSTM because it's presumably
+faster. We can cite some papers here.
+
+\section{Results}\label{results}
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\section{Discussion}\label{discussion}
 
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
+\section{References}\label{references}
+
+\section{Appendix}\label{appendix}
+
+Add computer code, plots, and other relevant technical details that will
+help me evaluating your work.
+
+\section*{Statement of contribution}\label{statement-of-contribution}
+\addcontentsline{toc}{section}{Statement of contribution}
 
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-QuickDra10}{}%
+\hypertarget{ref-QuickDra10}{}
 Google. 2018. ``Quick, Draw! Doodle Recognition Challenge \textbar{}

### 2018-12-03 proposal/proposal2.Rmd
organize

--- 2018-10-30 proposal.Rmd
@@
 Our team is basing our project idea on the "Quick, Draw!" Kaggle competition 
-which classifies user drawings into one of 340 label categories. Drawings may 
+which classifies user drawings into one of 345 label categories. Drawings may 
 be incomplete or not match the label. The challenge is to effectively build a 
 recognizer that can work with this noisy data. The challenge is unique from an
-algorithmic perspective because the data includes both time series and pattern
-recognition components; areas that have traditionally used two different 
-types of Neural Network Architectures. Our team uses a subset of the data to 
-provide a solution to the Kaggle problem while taking a more in depth 
-look at potential pitfalls and modeling strategy without having to use 
+algorithmic perspective because the data includes both temporal and 
+spatial components; areas that have traditionally used two different 
+types of Neural Network Architectures. Our team attempts to provide a solution 
+using a subset of the data, to the Kaggle problem, while taking a more in depth 
+look at potential pitfalls and modeling strategy, without having to use 
 extensive computational resources.
@@
 as the requested drawing category and country code. The format of each drawing
-array contains 3-dimensional vectors where components are the x,y coordinates
-and time in milliseconds since the first point. We use a preprocessed version
-of the dataset which is simplified and takes up about a tenth of the memory.
+array contains 3-dimensional vectors of strokes, where components are the x,y 
+coordinates and time in milliseconds since the first point. We use a preprocessed 
+version of the dataset which is simplified and is about tenth the size.
 The simplified dataset contains a CSV file for each of the 345 categories.
@@
 
-We plan to use a Neural Network for this problem. The problem is unique in
-that it has a time series component related to drawing stoke sequence, and
-a image recognition component related the cumulative strokes. Recurrent Neural
+We plan to use Neural Networks for this problem. The problem is unique in
+that it has a time series component related to drawing stroke sequences, and
+an image recognition component related to the cumulative strokes. Recurrent Neural
 Networks (RNN) are generally associated with time series. Convolutional Neural
@@
 time series and pattern recognition components requires us to consider 
-tradeoffs related to model accuracy, complexity, and algorithmic scalability.
+tradeoffs related to model accuracy, scalability, and algorithmic complexity.
 For example, RNNs have been shown to be less scalable than CNNs due to
@@
 attempted to capture benefits of RNNs in CNNs through the use of techniques
-such as dilation. We are still working through the best Neural Network 
-Architecture for this problem.
+such as dilation. The goal of this project is to compare the aforementioned
+Neural Network Architectures, and present the findings.
 
-Feature engineering is also an important part of our modeling strategy. We
+Feature engineering will help us capture the temporal component of our data. We
 recognize that choosing appropriate features can reduce the amount of data 

### 2018-12-05 presentation/final.Rmd
started putting together the presentation

---
title: "Quick, Draw! Challenge"
subtitle: DS 5220
author: "Blue team"
output:
  revealjs::revealjs_presentation:
    theme: simple
    highlight: pygments
    center: true
---

## Problem to Solve: Quick, Draw! Image Classification

* We focus on a subset of the Kaggle competition
  - Image classification vs. identifying sequences when drawing

## Our Strategy

* Bias vs. Variance
  - Overfitting 
  - Underfitting
* Multi-Class Classification
  - Accurcy vs. Number of Classes
  - Imbalanced classes
* Neural Network Architecture
  - Fully Connected Neural Network
  - Convolutional Neural Network (CNN)

# Bias vs. Variance

## Seeing the Big Picture

* Neural Networks differ from many other models in
  that bias vs. variance is less of a tradeoff

  - We can fix bias, then separately fix variance

## Overfitting

* overfitting is bad and here's how it happens
* here's how we dealt with it

## Underfitting

* also bad, here's how we deal with it

# Multi-Class Classification

## Why is this necessary?

* We need to classify multiple quickly drawn images

## Accurcy vs. Number of Classes

* Accuracy seemed to go down as the number 
  of classes went up

## Imbalanced classes

* Also bad, used under-sampling

# Neural Network Architecture

## Fully Connected Neural Network

* Designed using Tensorflow without Keras

## Convolutional Neural Network (CNN)

* Designed using Keras with Tensorflow backend

# Summary

## Recommendations

* Neural networks are good for classifying images containing a hotdog

# Thank you! Any questions?

### 2018-12-05 report/report_draft.Rmd
add tasks
report started

@@
 
+Provide a short background of the project (e.g., what kind of question is to be 
+answered with this work). Provide a short non-technical summary of your 
+analytical strategy.
+
 Our team has based our project idea on the "Quick, Draw!" Kaggle competition 
@@
 look at potential pitfalls and modeling strategy, without having to use 
-extensive computational resources.
+extensive computational resources. With this project, we explore image 
+classification in depth. We try out various experiments like fully connected 
+neural network vs CNN, bias vs. variance tradeoff, overfitting data purposefully
+and trying out techniques like regularization and dropout to fix overfitting and
+finally, testing neural network with imbalanced data.
 
@@
 
-First, we got our data from Google/Kaggle. CNNs are traditionally known to work
-well with spatial data, which prompted us to process the original csv files to 
-images of the drawings.
+First, let's talk about data collection. Google initially compressed and converted
+the images to .npy files. These .npy files are basically 28*28 images, containing
+more than 100,000 rows each.
 
-Maybe we should structure our models as experiments. Then talk 
-about what worked the best. Here's some example hypotheses
+So here are the experiments we performed.
+
+### EDA
+
+We checked the We plotted images from several categories
 
@@
 
+Try out: Overfitting, Bias/variance, imbalance, dropout, pushing activation
+
 # Results
@@
 
+The model built using CNN was better than human accuracy, which was in turn 
+better than the model built using fully connected network.
+
 # References
@@
 # Statement of contribution
+
+Everyone worked on everything.

### 2018-12-05 report/report_draft.tex
add tasks
report started

@@
 
+Provide a short background of the project (e.g., what kind of question
+is to be answered with this work). Provide a short non-technical summary
+of your analytical strategy.
+
 Our team has based our project idea on the ``Quick, Draw!'' Kaggle
@@
 in depth look at potential pitfalls and modeling strategy, without
-having to use extensive computational resources.
+having to use extensive computational resources. With this project, we
+explore image classification in depth. We try out various experiments
+like fully connected neural network vs CNN, bias vs.~variance tradeoff,
+overfitting data purposefully and trying out techniques like
+regularization and dropout to fix overfitting and finally, testing
+neural network with imbalanced data.
 
@@
 
-First, we got our data from Google/Kaggle. CNNs are traditionally known
-to work well with spatial data, which prompted us to process the
-original csv files to images of the drawings.
+First, let's talk about data collection. Google initially compressed and
+converted the images to .npy files. These .npy files are basically 28*28
+images, containing more than 100,000 rows each.
 
-Maybe we should structure our models as experiments. Then talk about
-what worked the best. Here's some example hypotheses
+So here are the experiments we performed.
+
+\subsubsection{EDA}\label{eda}
+
+We checked the We plotted images from several categories
 
@@
 faster. We can cite some papers here.
+
+Try out: Overfitting, Bias/variance, imbalance, dropout, pushing
+activation
 

### 2018-12-06 report/report_draft.Rmd
minor suggestions on report format

@@
 
-$H_a:$ Prediction will go up when we use a CNN compared to a fully
+$H1_a:$ Prediction will go up when we use a CNN compared to a fully
 	connected feedforward neural network.
 
-$H_a:$ Prediction will go up when we use regularization.
+$H2_a:$ Prediction will go up when we use regularization.
 
-$H_a:$ Run neural networks without any activations.
+$H3_a:$ Run neural networks without any activations.
 
-$H_a:$ Try different architectures- number of layers, different (or no) 
+$H4_a:$ Try different architectures- number of layers, different (or no) 
 activations
 
-$H_a:$ CNN with dilation is preferable to LSTM because it's presumably
+$H5_a:$ CNN with dilation is preferable to LSTM because it's presumably
 faster. We can cite some papers here.
@@
 hypotheses
+
+| Hypothesis | Result         |
+| ---------- | -------------- |
+| $H1_a$     | Accept         |
+| $H2_a$     | Fail to Accept |
+| $H3_a$     | Fail to Accept |
+| $H4_a$     | Accept         |
+| $H5_a$     | Fail to Accept |
+
+## Fully Connected Neural Network
+
+Here's the run down on it and how it relates
+to the hypotheses.
+
+## Convolutional Neural Network
+
+same thing for the CNN. TOOD: can we fold in some of
+the specializations into this section??
 

### 2018-12-06 report/report_draft.tex
minor suggestions on report format

@@
 \urlstyle{same}  % don't use monospace font for urls
+\usepackage{longtable,booktabs}
 \usepackage{graphicx,grffile}
@@
 
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 
-\section{Methods}\label{methods}
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
 
@@
 
-\subsubsection{EDA}\label{eda}
+\hypertarget{eda}{%
+\subsubsection{EDA}\label{eda}}
 
@@
 
-\(H_a:\) Prediction will go up when we use a CNN compared to a fully
+\(H1_a:\) Prediction will go up when we use a CNN compared to a fully
 connected feedforward neural network.
 
-\(H_a:\) Prediction will go up when we use regularization.
+\(H2_a:\) Prediction will go up when we use regularization.
 
-\(H_a:\) Run neural networks without any activations.
+\(H3_a:\) Run neural networks without any activations.
 
-\(H_a:\) Try different architectures- number of layers, different (or
+\(H4_a:\) Try different architectures- number of layers, different (or
 no) activations
 
-\(H_a:\) CNN with dilation is preferable to LSTM because it's presumably
-faster. We can cite some papers here.
+\(H5_a:\) CNN with dilation is preferable to LSTM because it's
+presumably faster. We can cite some papers here.
 
@@
 
-\section{Results}\label{results}
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
@@
 
-\section{Discussion}\label{discussion}
+\begin{longtable}[]{@{}ll@{}}
+\toprule
+Hypothesis & Result\tabularnewline
+\midrule
+\endhead
+\(H1_a\) & Accept\tabularnewline
+\(H2_a\) & Fail to Accept\tabularnewline
+\(H3_a\) & Fail to Accept\tabularnewline
+\(H4_a\) & Accept\tabularnewline
+\(H5_a\) & Fail to Accept\tabularnewline
+\bottomrule
+\end{longtable}
+
+\hypertarget{fully-connected-neural-network}{%
+\subsection{Fully Connected Neural
+Network}\label{fully-connected-neural-network}}
+
+Here's the run down on it and how it relates to the hypotheses.
+
+\hypertarget{convolutional-neural-network}{%
+\subsection{Convolutional Neural
+Network}\label{convolutional-neural-network}}
+
+same thing for the CNN. TOOD: can we fold in some of the specializations
+into this section??
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 
-\section{References}\label{references}
+The model built using CNN was better than human accuracy, which was in
+turn better than the model built using fully connected network.
 
-\section{Appendix}\label{appendix}
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+\hypertarget{appendix}{%
+\section{Appendix}\label{appendix}}
 
@@
 
-\section*{Statement of contribution}\label{statement-of-contribution}
-\addcontentsline{toc}{section}{Statement of contribution}
+\hypertarget{statement-of-contribution}{%
+\section{Statement of contribution}\label{statement-of-contribution}}
+
+Everyone worked on everything.
 
 \hypertarget{refs}{}
-\hypertarget{ref-QuickDra10}{}
+\leavevmode\hypertarget{ref-QuickDra10}{}%
 Google. 2018. ``Quick, Draw! Doodle Recognition Challenge \textbar{}

### 2018-12-07 report/draft.tex
changed wording and converted to latex

%
% PPUA 5262 Final
%
\documentclass[12pt]{article}

%
% Packages
%
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{clrscode3e}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage[table]{xcolor}% http://ctan.org/pkg/xcolor
\usepackage{caption}
\usepackage{subcaption}

%
% Package settings
%

%% begin.rcode setup, include=FALSE
% library(knitr)
%
% opts_chunk$set(fig.path='figure/latex-', cache.path='cache/latex-')
%% end.rcode

%
% Document Settings
%
\setlength{\parskip}{1pc}
\setlength{\parindent}{0pt}
\setlength{\topmargin}{-3pc}
\setlength{\textheight}{9.5in}
\setlength{\oddsidemargin}{0pc}
\setlength{\evensidemargin}{0pc}
\setlength{\textwidth}{6.5in}

\title{Quick, Draw! Image Classification}
\author{Group 3}

%
% Image directory location
%
\graphicspath{ {./imgs/} }

\begin{document}

\maketitle

\section{Introduction}

Many recent industry innovations have included neural networks. An
important research area within neural networks has been image classification.
One of the best techniques to classify images is Convolutional Neural
Networks (CNN). We have sought to better understand industry best practices
by exploring different image classification techniques with neural networks.
Our exploration starts by manually performing image classification and
noting the level of human accuracy. We then use a fully connected neural
network to perform image classification before implementing a CNN. We
implemented options for tuning and debugging the neural network while
implementing each method during our process. We were able to develop an
intuition for image classification best practices using neural networks
after our research.

Many supervised machine learning methods often differ from those employed 
when constructing neural networks. These supervised machine learning
methods, such as logistic regression or decision tree classification,
require feature engineering to successfully classify an image. Models
requiring feature engineering rely on considerable effort by subject matter
experts to achieve reasonable accuracy. Neural networks allow for an
approach which does not require feature engineering to achieve similar or
better accuracy. Industry engineering efforts regarding feature engineering
can be shifted to neural networks  in some cases such as image
classification. Moreover, the availability of cheap computing resources
sped up a shift towards the use of neural networks.

When considering modeling approaches, we examine both fully connected and 
convolutional neural networks. A fully connected neural network is known to 
be inefficient at classifying large images. These inefficiencies in a fully 
connected neural network partly arise because every neuron in the $\ell-1$
layer must be connected to the $\ell$th layer. This requires a series
of matrix multiplications and additions across each layer $\ell$.
Alternatively, a CNN allows for more efficient computations by: (1) efficient
and automatic feature generation using local connectivity and parameter
sharing of convolution operations, (2) dimensionality reduction using pooling
layers. In this project, we attempt to understand these topics in depth
from an empirical stanpoint in relation to image classification using
Google's ``Quick, Draw!'' dataset.

\section{Methods}

\subsection{EDA}

\section{Results}

\subsection{Image Classification using Humans}

\subsection{Fully Connected Neural Network}

\subsection{Convolutional Neural Network}

\section{Discussion}

\bibliographystyle{unsrt}
\bibliography{references}

\end{document}

### 2018-12-07 report/report_draft.Rmd
changed wording and converted to latex

@@
 ---
+
+# Introduction
+
+Many recent industry innovations have included neural networks. An important 
+research area within neural networks has been image classification. One of 
+the best techniques to classify images is Convolutional Neural Networks (CNN).
+We have sought to better understand industry best practices by exploring 
+different image classification techniques with neural networks. Our 
+exploration starts by manually performing image classification and noting the 
+level of human accuracy. We then use a fully connected neural network to 
+perform image classification before implementing a CNN. We implemented options
+for tuning and debugging the neural network while implementing each method 
+during our process. We were able to develop an intuition for image 
+classification best practices using neural networks after our research.
+
+Many supervised machine learning methods often differ from those employed 
+when constructing neural networks. These supervised machine learning methods, 
+such as logistic regression or decision tree classification, require feature 
+engineering to successfully classify an image. Models requiring feature 
+engineering rely on considerable effort by subject matter experts to achieve 
+reasonable accuracy. Neural networks allow for an approach which does not 
+require feature engineering to achieve similar or better accuracy. Industry 
+engineering efforts regarding feature engineering can be shifted to neural 
+networks  in some cases such as image classification. Moreover, the 
+availability of cheap computing resources sped up a shift towards the use of 
+neural networks.
+
+When considering modeling approaches, we consider both fully connected and 
+convolutional neural networks. A fully connected neural network is known to 
+be inefficient at classifying large images. These inefficiencies in a fully 
+connected neural network partly arise because every neuron in the $n-1$ layer 
+must be connected to the $n$th layer.
 

### 2018-12-07 report/draft.tex
moved report over to latex with minimal reformatting

@@
 
-\subsection{EDA}
+We generate a series of hypotheses to inform our analysis.
+
+\subsection{Hypotheses}
+
+We start off with hypothesis that we learnt from theoretical literature. We
+then perform several experiments to test the validity of these hypothesis and
+report out our findings. Some of our hypothesis is as follows:
+
+1) Fully connected neural network will struggle to get reasonable accuracy.
+In contrast CNN’s should perform better and faster than a Fully connected
+neural network.
+
+2) Training a deep neural network with limited data will result in
+overfitting. Conversely, adding more data to a deep neural network is an
+effective way to manage variance.
+
+3) Imbalance class prediction can be improved using data augmentation.
+
+4) Increasing the number of categories will cause a significant drop in the
+performance of the neural network. Increasing the overall data points for the
+network to be trained on should help.
+
+\subsection{Data}
+
+Google originally collected the data from users by explicitly asking them to
+draw objects like forks, hotdogs, etc. The original dataset is 70 GB in size
+but Google also provides a simplified version of the dataset- consisting only
+of the final images. We used these '.npy' files to train all of our models.
+These numpy bitmap files consist of more than 100,000 rows each and 784
+features, since they are basically $28 \times 28$ images. Samples of these
+images are shown in Figure \ref{fig:quickImages}. The subset of this
+dataset sampled varies from experiment to experiment. More details can be
+found in the experiments mentioned below.
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[scale=0.5]{fig1}
+  \end{center}
+  \caption{Sample images from Quick, Draw! dataset}
+  \label{fig:quickImages}
+  \end{figure}
 
@@
 
-\subsection{Image Classification using Humans}
+\subsection{Hypothesis 1}
 
-\subsection{Fully Connected Neural Network}
+Hypothesis 1: Fully connected neural network will struggle to get reasonable
+accuracy. In contrast CNN’s should perform better and faster than a Fully
+connected neural network.
 
-\subsection{Convolutional Neural Network}
+In order to perform this experiment we built a fully connected neural network
+using tensorflow. Our dataset sample selected for this experiment had 10
+categories: fish, fork, hotdog, flamingo, airplane, alarm clock, baseball bat,
+bicycle, dolphin, elephant. We had 100,000 examples in total, with 10,000
+example for each of the aforementioned categories. Our train:test split was
+chosen to be 80:20, an industry standard.
+
+Our initial objective was to build a model that overfit our given dataset,
+with the idea that once our model is complex enough to overfit the training
+data, we could regularize it in order to achieve a more optimal fit.
+
+After experimenting with various architectures, we narrowed down to a 3 layer
+model as specified in Figure \ref{fig:fcnn}.
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[scale=0.75]{fig2}
+    \end{center}
+  \caption{Fully Connected Neural Network Architecture}
+  \label{fig:fcnn}
+  \end{figure}
+
+The training set accuracy observed was 93\%, while the test accuracy was at
+73\%, which clearly indicated an overfit in the model, as designed.
+
+The model took 1500 epochs to converge taking around ~45 minutes on a machine with 13 GB of RAM, Intel Xeon CPU 2.3 GHz (1 core, 2 threads), 1xTesla K80 GPU having 2496 CUDA cores with 12 GB GDDR5 VRAM.
+In order to manage variance, we attempted regularizing using dropouts.
+Unfortunately, that effort was futile.
+We then doubled the amount of data sampled to train the model, eventually helping us achieve training set accuracy of 86\% and test set accuracy of 82\%.
+Our
+learning rate was specified as 0.001.
+
+While our initial hypothesis assumed that a regular fully connected network would struggle to achieve a reasonable accuracy, our experiment proved our hypothesis as invalid by achieving a commendable accuracy.
+Having said that, the model did take a longer to converge than expected.
+
+As for the CNN, we started off with the same approach with attempting to overfit the model using convolution operations and a fully connected layer.
+Since our objective is to categorise 10 categories, softmax was chosen as the activation function for the final layer, while every other neuron had ReLU as the activation function owing to ReLU’s faster convergence.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[scale=0.5]{fig3}
+    \end{center}
+  \caption{CNN Architecture}
+  \label{fig:cnnArch}
+\end{figure}
+
+Loss Function: Categorical Cross Entropy [TO DO: Why?]
+Optimizer: [TODO: Why?]
+Number of Trainable Parameters: 320,890
+
+The number of clearly being much greater than the number of observations used to train the model, we need to reduce the number of parameters vis-a-vis number of observations.
+From the available strategies to do so, we incorporated MaxPooling [TO DO: why?] post each convolution operation, in order to extract only the most significant features.
+As as a consequence, this reduced the volume of the output, giving the model a smaller memory footprint as well as a smaller number of trainable parameters.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[scale=0.5]{fig4}
+  \end{center}
+  \caption{CNN Architecture with Reduced Parameters}
+  \label{fig:cnnArchPool}
+\end{figure}
+
+The observable training accuracy is 0.90, whereas the validation accuracy is 0.89. The model converged after 10 epochs, which took ~ 3 minutes.
+
+Some of the other unsuccessful strategies used to decrease the number of parameters and to regularize the network were strides and dropouts which sacrificed way too much in the way of accuracy and thus had to be dropped.
+Since most of our drawings had most of their information in the center of the image, padding, in theory wouldn’t be of much help.
+However, actually implementing padding improved our accuracy by 1\% at a cost of ~6000 parameters.
+
+Conclusion
+
+A slight (~5\%) improvement in the accuracy of a CNN when compared to a fully connected network, keeping the number of parameters constant, was observed.
+However, the CNN happened to be around 13 times faster in practice.
+This delta in accuracy and speed is likely to be amplified as we scale the model up with respect to model complexity or data.
+
+\subsection{Hypothesis 2}
+
+Hypothesis 2: Training a deep neural network with limited data will result in overfitting. Conversely, adding more data to a deep neural network is an effective way to manage variance.
+
+With us reaching an optimal model for the dataset, we decided to study the effect of reduction in the sampled dataset and its relationship with overfitting.
+
+\subsection{Hypothesis 3}
+
+We tried to show the importance of having equal partitions of data to train a machine learning model. Imbalanced classes can make a model biased toward the majority class.
 

### 2018-12-07 report/draft.tex
tried to structure our report more

@@
 %
-% PPUA 5262 Final
+% DS 5220 Final
 %
@@
 
-We generate a series of hypotheses to inform our analysis.
+We generate hypotheses based on expectations related to model performance. We
+then conduct preprocessing steps on our data.
+
+\subsection{Overview of Statistical Modeling Techniques}
+
+We use human classification, fully connected neural networks,
+and a Convolutional Neural Network (CNN). We discuss the CNN in
+more detail than was covered in class so an overview is provided
+here.
+
+\subsubsection{Convolutional Neural Network (CNN)}
+
+TODO: explain how we used a CNN or anything else not covered in class
 
@@
 
-We start off with hypothesis that we learnt from theoretical literature. We
-then perform several experiments to test the validity of these hypothesis and
-report out our findings. Some of our hypothesis is as follows:
+We have generated the following hypotheses based on theoretical
+expectations:
 
-1) Fully connected neural network will struggle to get reasonable accuracy.
-In contrast CNN’s should perform better and faster than a Fully connected
-neural network.
+\subsubsection{Addressing model component of performance}
 
-2) Training a deep neural network with limited data will result in
-overfitting. Conversely, adding more data to a deep neural network is an
-effective way to manage variance.
+\begin{align*}
+  H1_a: &\text{Fully Connected Neural Network will struggle} \\
+  &\text{ to get reasonable accuracy.} \\
+  H2_a: &\text{CNN should perform better than fully} \\
+  &\text{connected neural network.} \\
+\end{align*}
 
-3) Imbalance class prediction can be improved using data augmentation.
+\subsubsection{Addressing data component of performance}
+\begin{align*}
+  H3_a: &\text{Small amounts of data will lead to overfitting.} \\
+  H4_a: &\text{Imbalanced class prediction can be improved}\\
+  &\text{using data augmentation.} \\
+  H5_a: &\text{Addition of a category to the trained model will} \\
+  &\text{reduce the performance of the model.} \\
+  H6_a: &\text{Adding more training data will improve the} \\
+  &\text{performance of the model.} \\
+\end{align*}
 
-4) Increasing the number of categories will cause a significant drop in the
-performance of the neural network. Increasing the overall data points for the
-network to be trained on should help.
+TODO: more context would be helpful here, maybe some citations
 
@@
 
-Google originally collected the data from users by explicitly asking them to
+Google originally collected the data from users by explicitly asking users to
 draw objects like forks, hotdogs, etc. The original dataset is 70 GB in size
@@
 
-\subsection{Hypothesis 1}
+We report our results by examining the validity of each hypothesis.
 
-Hypothesis 1: Fully connected neural network will struggle to get reasonable
-accuracy. In contrast CNN’s should perform better and faster than a Fully
-connected neural network.
+\subsection{Addressing model component of performance}
 
@@
 
-The model took 1500 epochs to converge taking around ~45 minutes on a machine with 13 GB of RAM, Intel Xeon CPU 2.3 GHz (1 core, 2 threads), 1xTesla K80 GPU having 2496 CUDA cores with 12 GB GDDR5 VRAM.
-In order to manage variance, we attempted regularizing using dropouts.
-Unfortunately, that effort was futile.
-We then doubled the amount of data sampled to train the model, eventually helping us achieve training set accuracy of 86\% and test set accuracy of 82\%.
-Our
-learning rate was specified as 0.001.
+The model took 1500 epochs to converge taking around ~45 minutes on a
+machine with 13 GB of RAM, Intel Xeon CPU 2.3 GHz (1 core, 2 threads),
+1xTesla K80 GPU having 2496 CUDA cores with 12 GB GDDR5 VRAM. In order to
+manage variance, we attempted regularizing using dropouts. Unfortunately,
+that effort was futile. We then doubled the amount of data sampled to train
+the model, eventually helping us achieve training set accuracy of 86\% and
+test set accuracy of 82\%. Our learning rate was specified as 0.001.
 
-While our initial hypothesis assumed that a regular fully connected network would struggle to achieve a reasonable accuracy, our experiment proved our hypothesis as invalid by achieving a commendable accuracy.
-Having said that, the model did take a longer to converge than expected.
+While our initial hypothesis assumed that a regular fully connected network
+would struggle to achieve a reasonable accuracy, our experiment proved our
+hypothesis as invalid by achieving a commendable accuracy. Having said that,
+the model did take a longer to converge than expected.
 
-As for the CNN, we started off with the same approach with attempting to overfit the model using convolution operations and a fully connected layer.
-Since our objective is to categorise 10 categories, softmax was chosen as the activation function for the final layer, while every other neuron had ReLU as the activation function owing to ReLU’s faster convergence.
+As for the CNN, we started off with the same approach with attempting to
+overfit the model using convolution operations and a fully connected layer.
+Since our objective is to categorise 10 categories, softmax was chosen as the
+activation function for the final layer, while every other neuron had ReLU as
+the activation function owing to ReLU’s faster convergence.
 
@@
 
-The number of clearly being much greater than the number of observations used to train the model, we need to reduce the number of parameters vis-a-vis number of observations.
-From the available strategies to do so, we incorporated MaxPooling [TO DO: why?] post each convolution operation, in order to extract only the most significant features.
-As as a consequence, this reduced the volume of the output, giving the model a smaller memory footprint as well as a smaller number of trainable parameters.
+The number of clearly being much greater than the number of observations used
+to train the model, we need to reduce the number of parameters vis-a-vis
+number of observations. From the available strategies to do so, we
+incorporated MaxPooling [TO DO: why?] post each convolution operation, in
+order to extract only the most significant features. As as a consequence,
+this reduced the volume of the output, giving the model a smaller memory
+footprint as well as a smaller number of trainable parameters.
 
@@
 
-The observable training accuracy is 0.90, whereas the validation accuracy is 0.89. The model converged after 10 epochs, which took ~ 3 minutes.
+The observable training accuracy is 0.90, whereas the validation accuracy is
+0.89. The model converged after 10 epochs, which took ~ 3 minutes.
 
-Some of the other unsuccessful strategies used to decrease the number of parameters and to regularize the network were strides and dropouts which sacrificed way too much in the way of accuracy and thus had to be dropped.
-Since most of our drawings had most of their information in the center of the image, padding, in theory wouldn’t be of much help.
-However, actually implementing padding improved our accuracy by 1\% at a cost of ~6000 parameters.
+Some of the other unsuccessful strategies used to decrease the number of
+parameters and to regularize the network were strides and dropouts which
+sacrificed way too much in the way of accuracy and thus had to be dropped.
+Since most of our drawings had most of their information in the center of the
+image, padding, in theory wouldn’t be of much help. However, actually
+implementing padding improved our accuracy by 1\% at a cost of ~6000
+parameters.
 
@@
 
-\subsection{Hypothesis 2}
+\subsection{Addressing data component of performance}
 
-Hypothesis 2: Training a deep neural network with limited data will result in overfitting. Conversely, adding more data to a deep neural network is an effective way to manage variance.
-
-With us reaching an optimal model for the dataset, we decided to study the effect of reduction in the sampled dataset and its relationship with overfitting.
-
-\subsection{Hypothesis 3}
-
-We tried to show the importance of having equal partitions of data to train a machine learning model. Imbalanced classes can make a model biased toward the majority class.
+ok
 
 \section{Discussion}
+
+Talk about our final slide and tie the hypotheses together
 

### 2018-12-07 report/draft.tex
see draft updates

@@
 
-We generate hypotheses based on expectations related to model performance. We
-then conduct preprocessing steps on our data.
+Our methodology consists of reviewing statistical methodology, presenting
+theoretically informed hypotheses, and providing an overview of our data.
 
@@
 
-\begin{align*}
-  H1_a: &\text{Fully Connected Neural Network will struggle} \\
-  &\text{ to get reasonable accuracy.} \\
-  H2_a: &\text{CNN should perform better than fully} \\
-  &\text{connected neural network.} \\
-\end{align*}
+\begin{itemize}
+  \item[$H1_a$] Fully Connected Neural Network will struggle to get reasonable
+    accuracy.
+  \item[$H2_a$] CNN should perform better than fully connected neural network.
+  \end{itemize}
 
 \subsubsection{Addressing data component of performance}
-\begin{align*}
-  H3_a: &\text{Small amounts of data will lead to overfitting.} \\
-  H4_a: &\text{Imbalanced class prediction can be improved}\\
-  &\text{using data augmentation.} \\
-  H5_a: &\text{Addition of a category to the trained model will} \\
-  &\text{reduce the performance of the model.} \\
-  H6_a: &\text{Adding more training data will improve the} \\
-  &\text{performance of the model.} \\
-\end{align*}
+
+\begin{itemize}
+\item[$H3_a$] Small amounts of data will lead to overfitting.
+\item[$H4_a$] Imbalanced class prediction can be improved using data
+  augmentation.
+\item[$H5_a$] Addition of a category to the trained model will reduce the
+  performance of the model.
+\item[$H6_a$] Adding more training data will improve the performance of the
+  model.
+\end{itemize}
 
@@
 
-Loss Function: Categorical Cross Entropy [TO DO: Why?]
-Optimizer: [TODO: Why?]
+\subsubsection{Understanding the Loss Function}
+
+Most modern neural networks are trained using maximum likelihood. This
+means that the cost function is the negative log-likelihood, also
+described as the cross-entropy between the training data and the model
+distribution \cite{Goodfellow-et-al-2016}. In our use case, we are interested
+in correctly classifying 10 categories of drawings. Our output unit
+use a Softmax function in our output layer because they are most often
+used for classifiers representing the probability distribution over
+$n$ different classes. Given features $h$, weights $W$, and bias $b$, a
+linear layer first predicts unnormalized log probabilities:
+\begin{equation}
+  z = W^Th + b 
+\end{equation}
+
+where $z_i = \log \tilde{P}(y = i | x)$. The softmax function can then
+exponentiate and normalize $z$ to obtain the desired $\hat{y}$. The
+softmax function is given by
+\begin{equation}
+  \func{softmax}(z)_i = \frac{\exp(z_i)}{\sum_j \exp(z_j)}.
+\end{equation}
+
+As a cost function, we wish to maximize
+$\log P(y=i; z) = \log \func{softmax}(z)_i$. We can do this formally as
+\begin{equation}
+  \log \func{softmax}(z)_i = z_i - \log \sum_j \exp(z_j).
+\end{equation}
+
+which gives us the intuition for a single example. Overall, we want to learn
+parameters that drive the softmax to predict the fraction of counts of each
+outcome observed in the training set \cite{Goodfellow-et-al-2016}:
+\begin{equation}
+  \func{softmax}(z(x;\theta))_i \approx \frac{\sum_{j=1}^m 1_{y^{(j)} = i,
+      x^{(j)} = x}}{\sum_{j=1}^m 1_{x^{(j)} = x}}.
+\end{equation}
+
+This is how the $\func{categorical\_crossetropy}(\text{y\_true}, \text{y\_pred})$ function is constructed in Keras \cite{chollet2015keras} as well as the
+$\func{softmax\_cross\_entropy\_with\_logits\_v2}(\text{labels}, \text{logits})$
+in Tensorflow \cite{tensorflow2015-whitepaper}.
+
+\subsubsection{Understanding the Optimization Technique}
+
+There are a number of different optimization methods such as Stochastic
+and Batch Gradient descent. In practice, we chose to use a faster variant
+of gradient descent called Adam. The Adam optimization is able to use
+an exponentially weighted average of the gradients, a technique called
+momentum, and then use that gradient to update your weights instead. We
+also want to slow down gradient descent in the horizontal direction, a
+technique called RMSprop), to reduce oscillations from batch gradient descent.
+The Adam optimization method is combining momentum and RMSprop to perform
+gradient descent more quickly \cite{DeepLear51:online}.
+
 Number of Trainable Parameters: 320,890
@@
 number of observations. From the available strategies to do so, we
-incorporated MaxPooling [TO DO: why?] post each convolution operation, in
+incorporated MaxPooling [TODO: why?] post each convolution operation, in
 order to extract only the most significant features. As as a consequence,

### 2018-12-07 report/draft.tex
updating section 2.1

@@
 We use human classification, fully connected neural networks,
-and a Convolutional Neural Network (CNN). We discuss the CNN in
-more detail than was covered in class so an overview is provided
-here.
+and a Convolutional Neural Network (CNN). We use more detail to
+describe techniques not covered during the course.
+
+\subsubsection{Understanding the Loss Function}
+
+Most modern neural networks are trained using maximum likelihood. This
+means that the cost function is the negative log-likelihood, also
+described as the cross-entropy between the training data and the model
+distribution \cite{Goodfellow-et-al-2016}. In our use case, we are interested
+in correctly classifying 10 categories of drawings. Our output unit
+use a Softmax function in our output layer because they are most often
+used for classifiers representing the probability distribution over
+$n$ different classes. Given features $h$, weights $W$, and bias $b$, a
+linear layer first predicts unnormalized log probabilities:
+\begin{equation}
+  z = W^Th + b 
+\end{equation}
+
+where $z_i = \log \tilde{P}(y = i | x)$. The softmax function can then
+exponentiate and normalize $z$ to obtain the desired $\hat{y}$. The
+softmax function is given by
+\begin{equation}
+  \func{softmax}(z)_i = \frac{\exp(z_i)}{\sum_j \exp(z_j)}.
+\end{equation}
+
+As a cost function, we wish to maximize
+$\log P(y=i; z) = \log \func{softmax}(z)_i$. We can do this formally as
+\begin{equation}
+  \log \func{softmax}(z)_i = z_i - \log \sum_j \exp(z_j).
+\end{equation}
+
+which gives us the intuition for a single example. Overall, we want to learn
+parameters that drive the softmax to predict the fraction of counts of each
+outcome observed in the training set \cite{Goodfellow-et-al-2016}:
+\begin{equation}
+  \func{softmax}(z(x;\theta))_i \approx \frac{\sum_{j=1}^m 1_{y^{(j)} = i,
+      x^{(j)} = x}}{\sum_{j=1}^m 1_{x^{(j)} = x}}.
+\end{equation}
+
+This is how the $\func{categorical\_crossetropy}(\text{y\_true}, \text{y\_pred})$ function is constructed in Keras \cite{chollet2015keras} as well as the
+$\func{softmax\_cross\_entropy\_with\_logits\_v2}(\text{labels}, \text{logits})$
+in Tensorflow \cite{tensorflow2015-whitepaper}.
+
+\subsubsection{Understanding the Optimization Technique}
+
+This course discussed optimization algorithms such as Stochastic Gradient
+Descent ($m =1$) and Batch Gradient Descent ($m=n \text{ obs.}$). Current
+best practices for Neural Networks recommend Mini-Batch Gradient Descent
+($m = c \text{ batch size}$). In practice, we use the Adam optimization
+algorithm due to faster computation time. The Adam optimization algorithm
+is able to achieve a speedup due to its combination of two other techniques.
+
+The Adam optimization algorithm takes an exponentially weighted average of
+the gradients, a technique called momentum, and then uses that gradient to
+update your weights. Secondly, we also want to slow down gradient descent in
+the horizontal direction with a technique called RMSprop to reduce
+oscillations from batch gradient descent. The Adam optimization method is
+combining momentum and RMSprop to perform gradient descent more quickly
+\cite{DeepLear51:online}.
 
@@
 Google originally collected the data from users by explicitly asking users to
-draw objects like forks, hotdogs, etc. The original dataset is 70 GB in size
-but Google also provides a simplified version of the dataset- consisting only
-of the final images. We used these '.npy' files to train all of our models.
-These numpy bitmap files consist of more than 100,000 rows each and 784
+draw objects like forks, hotdogs, etc \cite{QuickDra91:online}. The original
+dataset is 70 GB in size but Google also provides a simplified version of the
+dataset- consisting only of the final images \cite{googlecr61:online}. We
+used these '.npy' files to train all of our models. These numpy bitmap files
+consist of more than 100,000 rows each and 784
 features, since they are basically $28 \times 28$ images. Samples of these
@@
 
+Our analysis required us to use a subset of this data. Our dataset sample
+selected for this experiment had 10 categories: fish, fork, hotdog, flamingo,
+airplane, alarm clock, baseball bat, bicycle, dolphin, elephant. We had
+100,000 examples in total, with 10,000 example for each of the aforementioned
+categories.
+
 \section{Results}
@@
 In order to perform this experiment we built a fully connected neural network
-using tensorflow. Our dataset sample selected for this experiment had 10
-categories: fish, fork, hotdog, flamingo, airplane, alarm clock, baseball bat,
-bicycle, dolphin, elephant. We had 100,000 examples in total, with 10,000
-example for each of the aforementioned categories. Our train:test split was
-chosen to be 80:20, an industry standard.
+using tensorflow. Our train:test split was chosen to be 80:20, an industry
+standard.
 
@@
 \end{figure}
-
-\subsubsection{Understanding the Loss Function}
-
-Most modern neural networks are trained using maximum likelihood. This
-means that the cost function is the negative log-likelihood, also
-described as the cross-entropy between the training data and the model
-distribution \cite{Goodfellow-et-al-2016}. In our use case, we are interested
-in correctly classifying 10 categories of drawings. Our output unit
-use a Softmax function in our output layer because they are most often
-used for classifiers representing the probability distribution over
-$n$ different classes. Given features $h$, weights $W$, and bias $b$, a
-linear layer first predicts unnormalized log probabilities:
-\begin{equation}
-  z = W^Th + b 
-\end{equation}
-
-where $z_i = \log \tilde{P}(y = i | x)$. The softmax function can then
-exponentiate and normalize $z$ to obtain the desired $\hat{y}$. The
-softmax function is given by
-\begin{equation}
-  \func{softmax}(z)_i = \frac{\exp(z_i)}{\sum_j \exp(z_j)}.
-\end{equation}
-
-As a cost function, we wish to maximize
-$\log P(y=i; z) = \log \func{softmax}(z)_i$. We can do this formally as
-\begin{equation}
-  \log \func{softmax}(z)_i = z_i - \log \sum_j \exp(z_j).
-\end{equation}
-
-which gives us the intuition for a single example. Overall, we want to learn
-parameters that drive the softmax to predict the fraction of counts of each
-outcome observed in the training set \cite{Goodfellow-et-al-2016}:
-\begin{equation}
-  \func{softmax}(z(x;\theta))_i \approx \frac{\sum_{j=1}^m 1_{y^{(j)} = i,
-      x^{(j)} = x}}{\sum_{j=1}^m 1_{x^{(j)} = x}}.
-\end{equation}
-
-This is how the $\func{categorical\_crossetropy}(\text{y\_true}, \text{y\_pred})$ function is constructed in Keras \cite{chollet2015keras} as well as the
-$\func{softmax\_cross\_entropy\_with\_logits\_v2}(\text{labels}, \text{logits})$
-in Tensorflow \cite{tensorflow2015-whitepaper}.
-
-\subsubsection{Understanding the Optimization Technique}
-
-There are a number of different optimization methods such as Stochastic
-and Batch Gradient descent. In practice, we chose to use a faster variant
-of gradient descent called Adam. The Adam optimization is able to use
-an exponentially weighted average of the gradients, a technique called
-momentum, and then use that gradient to update your weights instead. We
-also want to slow down gradient descent in the horizontal direction, a
-technique called RMSprop), to reduce oscillations from batch gradient descent.
-The Adam optimization method is combining momentum and RMSprop to perform
-gradient descent more quickly \cite{DeepLear51:online}.
 

### 2018-12-07 report/draft.tex
section 2.1 looks good to me

@@
 This is how the $\func{categorical\_crossetropy}(\text{y\_true}, \text{y\_pred})$ function is constructed in Keras \cite{chollet2015keras} as well as the
-$\func{softmax\_cross\_entropy\_with\_logits\_v2}(\text{labels}, \text{logits})$
-in Tensorflow \cite{tensorflow2015-whitepaper}.
+$\func{softmax\_cross\_entropy\_with\_logits\_v2}(\text{labels}, \text{logits})$ in Tensorflow \cite{tensorflow2015-whitepaper}.
 
@@
 
-TODO: explain how we used a CNN or anything else not covered in class
+Convolutional networks, also known as convolutional neural networks(CNNs),
+are a specialized kind of neural network for processing data that has a
+known grid-like topology. Convolutional networks have been tremendously
+successful in practical applications. The name “convolutional neural network”
+indicates that the network employs a mathematical operation called
+convolution. Convolution is a specialized kind of linear operation.
+Convolutional networks are simply neural networks that use convolution
+in place of general matrix multiplication in at least one of their layers. See
+Figure \ref{fig:CNN} in the Appendix for a visual representation of a CNN.
+
+CNN architectures are usually composed of the building blocks as described
+below:
+
+\begin{enumerate}
+\item \textbf{Convolution:} Convolution operation helps to extract features
+  from the input image. The ﬁrst argument to the convolution is often referred
+  to as the input , the second argument as the feature detector (kernel) and
+  the output as the feature map. The convolution operation takes the dot
+  product of feature map and the input to detect features. Since each
+  convolution operation only focuses on a part of an image enabling parameter
+  sharing which helps to reduce the number of free parameters compared to a
+  fully connected neural network, allowing the network to be deeper with
+  fewer parameters.
+
+  \begin{itemize}
+  \item \textbf{Strides:} Stride is the number of pixels jumps between
+    each convolution operation performed on the input matrix. When the stride
+    is 1 then we move the filters to 1 pixel at a time. Increasing the stride
+    parameter reduces the output volume of the convolution since the jumps
+    are larger.
+  \item \textbf{Padding:}Padding helps to detect features present at the
+    border by adding a dummy values around the edge. Since convolution
+    operations collect most information from the center of the image, adding
+    padding allows the border pixels to be weighed more during the convolution
+    operation.
+  \end{itemize}
+
+  See Figure \ref{fig:CNNStridePad} in the Appendix for an example of Stride
+  and Padding.
+
+\item \textbf{Pooling Layer:} Pooling layers section helps to reduce the
+  number of parameters when the images are too large. Spatial pooling
+  (subsampling or downsampling) reduces the dimensionality of each map,
+  retaining important information. These can be of different types:
+  (1) Max Pooling - takes the largest element (most important feature)
+  from the rectified feature map (2) Average Pooling - averages all the
+  pixel values (3) Sum Pooling. sum of all pixels values. See figure
+  \ref{fig:CNNPooling} for a visualization of a Pooling Layer.
+
+\item \textbf{Fully Connected Layer:} The layer we call as FC layer, we
+  flattened our matrix into vector and feed it into a fully connected layer
+  like neural network.
+  \end{enumerate}
 
@@
 
+\section{Appendix}
+
+\begin{figure}[h!]
+  \begin{center}
+    \includegraphics[scale=0.5]{fig5}
+    \end{center}
+  \caption{CNN Architecture Example}
+  \label{fig:CNN}
+\end{figure}
+
+\begin{figure}[h!]
+  \begin{center}
+    \includegraphics[scale=0.5]{fig6}
+    \end{center}
+  \caption{Convolution operation with stride=1, padding=None}
+  \label{fig:CNNStridePad}
+\end{figure}
+
+\begin{figure}[h!]
+  \begin{center}
+    \includegraphics[scale=0.5]{fig7}
+  \end{center}
+  \caption{Example of Max Pooling}
+  \label{fig:CNNPooling}
+  \end{figure}
+
 \end{document}

### 2018-12-07 report/draft.tex
updating section 2.4

@@
 We have generated the following hypotheses based on theoretical
-expectations:
+expectations expressed in the literature \cite{Goodfellow-et-al-2016,
+  DeepLear51:online}. Performance is effected by both the model and
+data \cite{DeepLear51:online} so our hypotheses pertain to both effects.
 
@@
 \end{itemize}
-
-TODO: more context would be helpful here, maybe some citations
 
@@
 
+\subsection{Modeling Strategy}
+
+We provide our modeling strategy as it relates to the exploration of
+our stated hypotheses.
+
+\subsubsection{Baseline image classification using humans}
+
+Two group members completed a Human Intelligence Task by manually
+classifying images into one of our ten defined categories. This
+approach helped us get a benchmark which we could use to
+define ``reasonable accuracy''.
+
+\subsubsection{$H1_a$: Fully connected neural network will struggle
+  to get reasonable accuracy}
+
+In order to perform this experiment, we built a fully connected neural network
+using Tensorflow. Our train:test split was chosen to be 80:20, an industry
+standard. Our initial objective was to build a model that overfit our given
+dataset, with the idea that once our model is complex enough to overfit the
+training data, we could regularize it in order to achieve a more optimal fit.
+The fully connected neural network had 20,067 parameters.
+
+After experimenting with various architectures, we narrowed down to a 3 layer
+model as specified in Figure \ref{fig:fcnn} in the Appendix. The model was
+run on a machine with 13 GB of RAM, Intel Xeon CPU 2.3 GHz (1 core, 2
+threads), 1xTesla K80 GPU having 2496 CUDA cores with 12 GB GDDR5 VRAM.
+
+\subsubsection{$H2_a$ CNN should perform better than fully connected neural
+  network}
+
+As for the CNN, we started off with the same approach with attempting
+to overfit the model using convolution operations and a fully connected
+layer with an objective to categorise 10 categories, softmax was chosen
+as the activation function for the final layer, while every other neuron
+had ReLU as the activation function owing to ReLU’s faster convergence. See
+Figure \ref{fig:cnnArch} for a representation of this first CNN model.
+
+The number of parameters is much greater than the number of observations
+used to train the model resulting in overfitting. In order to reduce the
+number of parameters, we incorporated Pooling (Max Pooling) post each
+convolution operation extracting only the most significant features
+resulting in reduction of the volume of the convolutional output, giving
+the model a smaller memory footprint as well as a smaller number of
+trainable parameters. Figure \ref{fig:cnnArchPool} has a representation of this
+second CNN model. We employ strategies to reduce overfitting related to
+dropouts and stride in addition to adding more training data.
+
 \section{Results}
@@
 model as specified in Figure \ref{fig:fcnn}.
-
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[scale=0.75]{fig2}
-    \end{center}
-  \caption{Fully Connected Neural Network Architecture}
-  \label{fig:fcnn}
-  \end{figure}
 
@@
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[scale=0.5]{fig3}
-    \end{center}
-  \caption{CNN Architecture}
-  \label{fig:cnnArch}
-\end{figure}
-
 Number of Trainable Parameters: 320,890
@@
 footprint as well as a smaller number of trainable parameters.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[scale=0.5]{fig4}
-  \end{center}
-  \caption{CNN Architecture with Reduced Parameters}
-  \label{fig:cnnArchPool}
-\end{figure}
 
@@
   \begin{center}
+    \includegraphics[scale=0.75]{fig2}
+    \end{center}
+  \caption{Fully Connected Neural Network Architecture}
+  \label{fig:fcnn}
+\end{figure}
+
+\begin{figure}[h!]
+  \begin{center}
+    \includegraphics[scale=0.5]{fig3}
+    \end{center}
+  \caption{(CNN Model 1) Loss Function: Categorical Cross Entropy, Optimizer:
+    Adam, Trainable Parameters: 320,890}
+  \label{fig:cnnArch}
+\end{figure}
+
+\begin{figure}[h!]
+  \begin{center}
+    \includegraphics[scale=0.5]{fig4}
+  \end{center}
+  \caption{(CNN Model 2) Loss Function: Categorical Cross Entropy, Optimizer:
+    Adam, Trainable Parameters: 19,322}
+  \label{fig:cnnArchPool}
+\end{figure}
+
+\begin{figure}[h!]
+  \begin{center}
     \includegraphics[scale=0.5]{fig5}

### 2018-12-07 report/draft.tex
updating results section

@@
 %
-\documentclass[12pt]{article}
+\documentclass[11pt]{article}
 
@@
 
-\subsubsection{$H2_a$ CNN should perform better than fully connected neural
+\subsubsection{$H2_a$: CNN should perform better than fully connected neural
   network}
@@
 
+\subsubsection{$H3_a$: Small amounts of data will lead to overfitting}
+
+Once we have found an optimal model for the dataset, we use this model
+to study the effect of reduction in the sampled data. The optimal model
+helps inform the relative performance of models trained with lower amounts
+of data.
+
+\subsubsection{$H4_a$: Imbalanced class prediction can be improved using data
+  augmentation}
+
+We tried to show the importance of having equal partitions of data to train
+a machine learning model. Imbalanced classes can make a model biased toward
+the majority class. We trained the neural network twice, on an imbalanced
+dataset, keeping the number of samples for the majority class (fork), fixed
+and decreasing the number of samples for the minority class (hammer).
+
+The dataset was split into 80\% and 20\% for train and test respectively.
+This split had 0.2\%, 2\%, 10\%, 20\%, 50\%, and 100\% of majority class
+observations for the minority class. We designed the test dataset with 20000
+fresh samples from the main dataset to get a clearer picture of the model
+performance and we found that the accuracy on this fresh test dataset kept
+on increasing as the number of samples in the minority class increased.
+
+\subsubsection{$H5_a$ \& $H6_a$: Addition of a category to the trained model will reduce the performance of the model \& Adding more training data will improve the performance of the model}
+
+We try and understand the limits and capacity of the finalized model by
+varying the number of categories and the number of rows per category. We
+then measure the response accuracy and response loss of the model. The
+number of rows per category is altered in steps of 1000 from 7000 to
+20000 and the number of categories is altered in steps of 1 from 3 to 10
+categories. We attempt to formalize the patterns we see from visualizing
+the data with an OLS regression.
+
 \section{Results}
@@
 
-In order to perform this experiment we built a fully connected neural network
-using tensorflow. Our train:test split was chosen to be 80:20, an industry
-standard.
+\subsubsection{$H1_a$\: Fully Connected Neural Network will struggle to
+  get reasonable accuracy}
 
-Our initial objective was to build a model that overfit our given dataset,
-with the idea that once our model is complex enough to overfit the training
-data, we could regularize it in order to achieve a more optimal fit.
+While our initial hypothesis assumed that a regular fully connected
+network would struggle to achieve a reasonable accuracy, our experiment
+proved our hypothesis as invalid by achieving a commendable accuracy.
+Having said that, the model did take a longer to converge than expected.
 
-After experimenting with various architectures, we narrowed down to a 3 layer
-model as specified in Figure \ref{fig:fcnn}.
+\subsubsection{$H2_a$\: CNN should perform better than fully connected neural
+  network}
 
-The training set accuracy observed was 93\%, while the test accuracy was at
-73\%, which clearly indicated an overfit in the model, as designed.
-
-The model took 1500 epochs to converge taking around ~45 minutes on a
-machine with 13 GB of RAM, Intel Xeon CPU 2.3 GHz (1 core, 2 threads),
-1xTesla K80 GPU having 2496 CUDA cores with 12 GB GDDR5 VRAM. In order to
-manage variance, we attempted regularizing using dropouts. Unfortunately,
-that effort was futile. We then doubled the amount of data sampled to train
-the model, eventually helping us achieve training set accuracy of 86\% and
-test set accuracy of 82\%. Our learning rate was specified as 0.001.
-
-While our initial hypothesis assumed that a regular fully connected network
-would struggle to achieve a reasonable accuracy, our experiment proved our
-hypothesis as invalid by achieving a commendable accuracy. Having said that,
-the model did take a longer to converge than expected.
-
-As for the CNN, we started off with the same approach with attempting to
-overfit the model using convolution operations and a fully connected layer.
-Since our objective is to categorise 10 categories, softmax was chosen as the
-activation function for the final layer, while every other neuron had ReLU as
-the activation function owing to ReLU’s faster convergence.
-
-Number of Trainable Parameters: 320,890
-
-The number of clearly being much greater than the number of observations used
-to train the model, we need to reduce the number of parameters vis-a-vis
-number of observations. From the available strategies to do so, we
-incorporated MaxPooling [TODO: why?] post each convolution operation, in
-order to extract only the most significant features. As as a consequence,
-this reduced the volume of the output, giving the model a smaller memory
-footprint as well as a smaller number of trainable parameters.
-
-The observable training accuracy is 0.90, whereas the validation accuracy is
-0.89. The model converged after 10 epochs, which took ~ 3 minutes.
-
-Some of the other unsuccessful strategies used to decrease the number of
-parameters and to regularize the network were strides and dropouts which
-sacrificed way too much in the way of accuracy and thus had to be dropped.
-Since most of our drawings had most of their information in the center of the
-image, padding, in theory wouldn’t be of much help. However, actually
-implementing padding improved our accuracy by 1\% at a cost of ~6000
-parameters.
-
-Conclusion
-
-A slight (~5\%) improvement in the accuracy of a CNN when compared to a fully connected network, keeping the number of parameters constant, was observed.
-However, the CNN happened to be around 13 times faster in practice.
-This delta in accuracy and speed is likely to be amplified as we scale the model up with respect to model complexity or data.
+A slight (~5\%) improvement in the accuracy of a CNN when compared to a fully
+connected network, keeping the number of parameters constant, was observed.
+However, the CNN happened to be around 13 times faster in practice. This
+delta in accuracy and speed is likely to be amplified as we scale the model
+up with respect to model complexity or data.
 
@@
 
-ok
+\subsubsection{$H3_a$\: Small amounts of data will lead to overfitting}
+
+As found in the experiments, after decreasing the amount of data resulted
+in overfitting thereby confirming our hypothesis.
+
+\subsubsection{$H4_a$\: Imbalanced class prediction can be improved using
+  data augmentation}
+
+We get unexpectedly high accuracy of 43\% even in the worst case of 1:500
+imbalance in the categories while attempting to classify only the minority
+categories. This is because neural networks tend to generalize well on the
+most important weights.
+
+\subsubsection{$H5_a$ \& $H6_a$\: Addition of a category to the trained model
+  will reduce the performance of the model/Adding more training data will
+  improve the performance of the model}
+
+It can be interpreted from the regression that for every additional category,
+we can expect the test accuracy go down by 1.88\% and test loss to go up by
+6.4\% while for every additional datapoint, we can expect the accuracy to go
+up by .0016\% and loss to go down by .0056\%. Both results are statistically
+significant with reasonably tight confidence intervals and standard errors.
 
@@
 
-\begin{figure}[h!]
-  \begin{center}
-    \includegraphics[scale=0.75]{fig2}
-    \end{center}
-  \caption{Fully Connected Neural Network Architecture}
-  \label{fig:fcnn}
-\end{figure}
-
-\begin{figure}[h!]
-  \begin{center}
-    \includegraphics[scale=0.5]{fig3}
-    \end{center}
-  \caption{(CNN Model 1) Loss Function: Categorical Cross Entropy, Optimizer:
-    Adam, Trainable Parameters: 320,890}
-  \label{fig:cnnArch}
-\end{figure}
-
-\begin{figure}[h!]
-  \begin{center}
-    \includegraphics[scale=0.5]{fig4}
-  \end{center}
-  \caption{(CNN Model 2) Loss Function: Categorical Cross Entropy, Optimizer:
-    Adam, Trainable Parameters: 19,322}
-  \label{fig:cnnArchPool}
-\end{figure}
+% Stats summary figures
 
@@
 
+% Fully Connected Neural Network specs
+
+\begin{figure}[h!]
+  \begin{center}
+    \includegraphics[scale=0.75]{fig2}
+    \end{center}
+  \caption{Fully Connected Neural Network Architecture}
+  \label{fig:fcnn}
+\end{figure}
+
+% CNN model specs
+
+\begin{figure}[h!]
+  \begin{center}
+    \includegraphics[scale=0.5]{fig3}
+    \end{center}
+  \caption{(CNN Model 1) Loss Function: Categorical Cross Entropy, Optimizer:
+    Adam, Trainable Parameters: 320,890}
+  \label{fig:cnnArch}
+\end{figure}
+
+\begin{figure}[h!]
+  \begin{center}
+    \includegraphics[scale=0.5]{fig4}
+  \end{center}
+  \caption{(CNN Model 2) Loss Function: Categorical Cross Entropy, Optimizer:
+    Adam, Trainable Parameters: 19,322}
+  \label{fig:cnnArchPool}
+\end{figure}
+
+% H3 results
+
+\begin{figure}[h!]
+  \begin{center}
+    \includegraphics[scale=0.5]{fig8}
+    \end{center}
+  \caption{$H3_a$ Results for various size training data}
+  \label{fig:h3Results}
+  \end{figure}
+
+% H4 results
+
+\begin{figure}[h!]
+  \begin{center}
+    \includegraphics[scale=0.5]{fig9}
+    \end{center}
+  \caption{$H4_a$ Results for various class imbalances}
+  \label{fig:h4Results}
+\end{figure}
+
+% H5 results
+
+\begin{figure}[h!]
+  \begin{center}
+    \includegraphics[scale=0.5]{fig10}
+    \end{center}
+  \caption{$H5_a$ Results for various number of classes $k$}
+  \label{fig:h5Results}
+\end{figure}
+
+% H6 results
+
+\begin{figure}[h!]
+  \begin{center}
+    \includegraphics[scale=0.5]{fig11}
+    \end{center}
+  \caption{$H6_a$ Results for various amounts of training data $n$}
+  \label{fig:h6Results}
+  \end{figure}
+
 \end{document}

### 2018-12-07 report/draft.tex
please review the results section

@@
 
-We report our results by examining the validity of each hypothesis.
+We report our results by examining the validity of each hypothesis
+related to the model and data components of performance.
 
 \subsection{Addressing model component of performance}
+
+These results pertain to hypotheses effected by modeling choices.
 
@@
 
+These results pertain to hypotheses effected by the underlying
+distribution of data.
+
 \subsubsection{$H3_a$\: Small amounts of data will lead to overfitting}
@@
 As found in the experiments, after decreasing the amount of data resulted
-in overfitting thereby confirming our hypothesis.
+in overfitting thereby confirming our hypothesis. Figure \ref{fig:h3Results}
+contains a table with results for various number $n$ of observations.
+We increased the number of categories the Neural Network had to model,
+holding the number of examples per categories constant. Observations were
+increased in steps of 1,000 from $n=7000$ to $n=20,000$ examples. The
+number of categories $k$ ranged from $k=3$ to $k=10$ in steps of 1.
 
@@
 
+Results can be seen in the Appendix with Figure \ref{fig:h4Results}. 
 We get unexpectedly high accuracy of 43\% even in the worst case of 1:500
@@
 
-It can be interpreted from the regression that for every additional category,
+Results for various number of classes $k$ are visualized in the Appendix
+under Figure \ref{fig:h5Results}. Results for various amounts of training
+data are under Figure \ref{fig:h6Results} in the Appendix. It can be
+interpreted from the regression that for every additional category,
 we can expect the test accuracy go down by 1.88\% and test loss to go up by

### 2018-12-08 report/draft.tex
draft ready for review, let me know if the grammar is atrocious

@@
 
+\subsection{Baseline image classification using humans}
+
+Our baseline performance when manually classifying images into one
+of 10 categories is 87\% accuracy on average. We consider this accuracy to
+be ``reasonable accuracy''. The accuracy serves as a baseline for informing
+our understanding of results for our models.
+
 \subsection{Addressing model component of performance}
@@
 Results can be seen in the Appendix with Figure \ref{fig:h4Results}.
-We get unexpectedly high accuracy of 43\% even in the worst case of 1:500
-imbalance in the categories while attempting to classify only the minority
-categories. This is because neural networks tend to generalize well on the
-most important weights.
+Increasing the number of categories will cause a significant drop in the
+performance of the neural network. Increasing the overall data points for
+the network to be trained on should help.
+
+We try and understand the limits and capacity of the finalized model by
+varying the number of categories and the number of rows per category. We
+then measure the response accuracy and response loss of the model. The
+number of rows per category is altered in steps of 1000 from 7,000 to 20,000
+and the number of categories is altered in steps of 1 from 3 to 10.
+
+3 examples of the accuracy measurements with most rows (14,000), least rows
+(2,000) and intermediate rows (7,000) and their performance per every
+additional category is plotted below.
 
@@
 
-Talk about our final slide and tie the hypotheses together
+Our hypotheses explored how modeling choices and data distribution affect
+the accuracy performance. We found convolutional neural networks to be a
+good approach for classifying spatial data like images. Convolutional neural
+networks gave a better accuracy than humans which in turn was better than
+accuracy by fully connected neural networks.
+
+In terms of modeling choices, Fully connected neural network(FC-NN)
+performed reasonably well for the Quick Draw dataset classification. We
+noted that Convolutional Neural network’s(CNN) training and testing accuracy
+was 5 percent better than the FC-NN. Additionally,  CNN was 13 times faster
+to train than FC-NN. These results highlighted the optimizations done by
+CNN with local connectivity and parameter sharing. The change in accuracy
+and speed is likely to be amplified as we scale the model up with respect
+to model complexity or data.
+
+In terms of data distribution, we noted that for every additional category
+added to the finalized model architecture resulted in a 1.88\% decrease in
+test accuracy. Training the model on an imbalanced distribution of classes
+resulted in a poor prediction accuracy for the minority class. The
+prediction accuracy for the minority class increased as the imbalance in
+the training dataset decreased.
+
+In the future, we would like to make use of the time-series data for the
+drawings to predict the object while it is being drawn.
 
@@
 \section{Appendix}
+
+Our project code is available in a GitHub repo: ``https://github.com/tbonza/ds5220''.
 
@@
     \end{center}
-  \caption{$H3_a$ Results for various size training data}
+  \caption{$H3_a$ Results for various size training data at (epoch 1,
+    epoch 10)}
   \label{fig:h3Results}

### 2018-12-08 report/draft.tex
updating section 3.3.2

@@
 distribution \cite{Goodfellow-et-al-2016}. In our use case, we are interested
-in correctly classifying 10 categories of drawings. Our output unit
-use a Softmax function in our output layer because they are most often
-used for classifiers representing the probability distribution over
-$n$ different classes. Given features $h$, weights $W$, and bias $b$, a
+in correctly classifying 10 categories of drawings. Our output layer
+uses a Softmax function to represent the probability distribution over
+$k$ different classes. Given features $h$, weights $W$, and bias $b$, a
 linear layer first predicts unnormalized log probabilities:
@@
 \begin{enumerate}
-\item \textbf{Convolution:} Convolution operation helps to extract features
-  from the input image. The ﬁrst argument to the convolution is often referred
-  to as the input , the second argument as the feature detector (kernel) and
-  the output as the feature map. The convolution operation takes the dot
-  product of feature map and the input to detect features. Since each
-  convolution operation only focuses on a part of an image enabling parameter
-  sharing which helps to reduce the number of free parameters compared to a
-  fully connected neural network, allowing the network to be deeper with
-  fewer parameters.
+\item \textbf{Convolution:} The convolution operation helps to extract features
+  from the input image. The first argument is often referred
+  to as the input. The second is a feature detector (kernel) and
+  the output is the feature map. The convolution operation takes the dot
+  product of the feature detector and the input. Each convolution operation
+  focuses on a subset of an image. This approach allows for parameter
+  sharing which leads to a reduction in parameters relative to a
+  fully connected neural network. The advantage 
+  is that we can have a deeper network with fewer parameters.
 
@@
     are larger.
-  \item \textbf{Padding:}Padding helps to detect features present at the
+  \item \textbf{Padding:} Padding helps to detect features present at the
     border by adding a dummy values around the edge. Since convolution
@@
 
-\item \textbf{Fully Connected Layer:} The layer we call as FC layer, we
-  flattened our matrix into vector and feed it into a fully connected layer
-  like neural network.
+\item \textbf{Fully Connected (FC) Layer:} The FC layer flattens our
+  matrix into vector and feeds it into a fully connected neural network layer.
   \end{enumerate}
@@
 airplane, alarm clock, baseball bat, bicycle, dolphin, elephant. We had
-100,000 examples in total, with 10,000 example for each of the aforementioned
+100,000 examples in total, with 10,000 examples for each of the aforementioned
 categories.
@@
 
-As found in the experiments, after decreasing the amount of data resulted
+As found in the experiments, decreasing the amount of data resulted
 in overfitting thereby confirming our hypothesis. Figure \ref{fig:h3Results}
-contains a table with results for various number $n$ of observations.
-We increased the number of categories the Neural Network had to model,
-holding the number of examples per categories constant. Observations were
-increased in steps of 1,000 from $n=7000$ to $n=20,000$ examples. The
-number of categories $k$ ranged from $k=3$ to $k=10$ in steps of 1.
+contains a table with results for various number $n$ of observations. We
+can see in Figure \ref{fig:h3Results} that increasing the number of
+observations enables us tune a well-trained model.
 
@@
 
-Results can be seen in the Appendix with Figure \ref{fig:h4Results}.
-Increasing the number of categories will cause a significant drop in the
-performance of the neural network. Increasing the overall data points for
-the network to be trained on should help.
+We found that imbalanced class prediction can be improved by adding more
+data to the minority class. As shown in Figure \ref{fig:h4Results},
 
-We try and understand the limits and capacity of the finalized model by
-varying the number of categories and the number of rows per category. We
-then measure the response accuracy and response loss of the model. The
-number of rows per category is altered in steps of 1000 from 7,000 to 20,000
-and the number of categories is altered in steps of 1 from 3 to 10.
+shows us that
 
-3 examples of the accuracy measurements with most rows (14,000), least rows
-(2,000) and intermediate rows (7,000) and their performance per every
-additional category is plotted below.
+a
+decrease in class imbalance leads to an increase in prediction accuracy
+for the minority class. These predictions were based on 20,000 total
+samples which were previously unseen.
 

### 2018-12-08 report/draft.tex
review 3.3.2

@@
 
-We found that imbalanced class prediction can be improved by adding more
-data to the minority class. As shown in Figure \ref{fig:h4Results},
-
-shows us that
-
-a
-decrease in class imbalance leads to an increase in prediction accuracy
-for the minority class. These predictions were based on 20,000 total
-samples which were previously unseen.
+We are able to confirm that adding data to the minority class to reduce
+class imbalance increases prediction accuracy. Figure \ref{fig:h4Results}
+shows results that support this assertion by gradually decreasing the class
+imbalance by increasing the number of samples within the minority class.
+Minority class accuracy ranged from 43\% with a 1:500 class imbalance to
+98\% with a 1:1 class imbalance.These predictions were based on 20,000 total
+samples which were previously unseen. The neural network models we
+tested tended to generalize well on the most important weights, while
+attempting to classify only the minority categories.
 

### 2018-12-08 report/draft.tex
latest changes

@@
 
-A slight (~5\%) improvement in the accuracy of a CNN when compared to a fully
-connected network, keeping the number of parameters constant, was observed.
-However, the CNN happened to be around 13 times faster in practice. This
-delta in accuracy and speed is likely to be amplified as we scale the model
-up with respect to model complexity or data.
+We found a slight (~5\%) improvement in the accuracy of a CNN when compared
+to a similarly parameterized fully connected neural network. However, training
+time for the CNN was about 13 times faster in practice. The difference in
+accuracy and speed is likely to be amplified as we scale the model up in
+respect to complexity of amount of data.
 
@@
 98\% with a 1:1 class imbalance.These predictions were based on 20,000 total
-samples which were previously unseen. The neural network models we
-tested tended to generalize well on the most important weights, while
-attempting to classify only the minority categories.
+samples which were previously unseen. Neural network models tend to
+generalize well on the most important weights, while attempting to classify
+only the minority categories.
 
@@
 6.4\% while for every additional datapoint, we can expect the accuracy to go
-up by .0016\% and loss to go down by .0056\%. Both results are statistically
+up by 0.0016\% and loss to go down by 0.0056\%. Both results are statistically
 significant with reasonably tight confidence intervals and standard errors.
